---
title: "Swiftkey Data Exploration"
author: "Enrique Pérez Herrero"
date: "March 19, 2016"
email: eph.project1500@gmail.com
output: html_document
---

## Coursera Data Science Specialization Capstone: Milestone Report

## 0. EXECUTIVE SUMMARY.
This report is a milestone in the _Data Science Specialization Capstone_ from 
_Coursera_ by *Johns Hopkins University*.

The data is from a corpus called HC Corpora (www.corpora.heliohost.org). See the readme file at [http://www.corpora.heliohost.org/aboutcorpus.html](http://www.corpora.heliohost.org/aboutcorpus.html) for details on the corpora available.
The data came from a series of text documents in severallanguages from blogs, news and twitter content and came from [Swiftkey](https://swiftkey.com).

The report uses Natural Language Processing (NLP) techniques to explore the
data sets, the code includes several functions to easily expand the exploration
to other languages or other data sets. This milestone is intended to be used to
create a interactive app using [Shiny](http://shiny.rstudio.com/) package.

## 1. CLEANING ENVIRONMENT AND LOADING PACKAGES

```{r, message=FALSE, results="hide"}
rm(list = ls(all = TRUE))
gc(reset = TRUE)

library(tm)
library(readr)
library(wordcloud)
library(RWeka)
library(stringi)
library(reshape2)
library(ggplot2)
library(RColorBrewer)
library(knitr)
```

#### 1.1 LOADING DATA ACCESS FUNCTIONS

The functions

```{r}
source("Swiftkey_Data_Access_Functions.R")
```

#### 1.2 PARAMETERS

```{r}
AVAILABLE_LANGUAGES
LANGUAGE <- AVAILABLE_LANGUAGES[1]
LANGUAGE
SAMPLE_SIZE <- 1000
```

#### 1.3 LIST DATA FILES

```{r}
files_list <- get_files_path(LANGUAGE)
names(files_list) <- list.files(get_path(LANGUAGE), pattern = "*.txt")
```


### 2. FILES SUMMARY

```{r}
Mb <- 2**20 # One Megabyte
# Files names
files_name <- as.vector(names(files_list))

# Files sizes
files_size <- round(file.info(files_list)$size / Mb)

# Number of lines
files_lines <- sapply(files_list,
                      function(x) length(readLines(x, warn = FALSE)))
# Testing encoding
files_encoding <-
  sapply(files_list, function(x)
    guess_encoding(x, n_max = 1000))
files_encoding <- t(files_encoding)

# Counting words
files_stats <- sapply(files_list,
               function(x) stri_stats_latex(readLines(x, warn = FALSE)))

# Creating data frame.
files_words <- data.frame(t(files_stats))$Words
files_summary_df <- data.frame(row.names = files_name,
                               files_size,
                               files_lines,
                               files_words,
                               files_encoding)
files_summary_df <- t(files_summary_df)
```

```{r}
print(files_summary_df)
```


### 3. SAMPLE TEXT FILES

```{r}
set.seed(pi)
sapply(get_files_path(LANGUAGE),
       function(x) sample_text_file(x, size = SAMPLE_SIZE)
       )

# Directory for sample files.
sample_dir(LANGUAGE)
```

### 4. CREATE SAMPLE FILES CORPUS

```{r}
crps <- Corpus(DirSource(sample_dir(LANGUAGE), encoding = "UTF-8"),
              readerControl = list(language = LANGUAGE))
```

#### 4.1 Check corpus

```{r}
inspect(crps)
meta(crps, "id")
```


### 5. CORPUS TRANSFORMATIONS

```{r}
corpus_cleaning <- function(my_corpus){
  my_corpus <- tm_map(my_corpus, stripWhitespace)
  my_corpus <- tm_map(my_corpus, removePunctuation)
  my_corpus <- tm_map(my_corpus, content_transformer(tolower))
  my_corpus <- tm_map(my_corpus, removeNumbers)
  return(my_corpus)
}

crps <- corpus_cleaning(crps)
```


### 6. DISTRIBUTION OF WORD FREQUENCIES

```{r}
ngram_Tokenizer <- function(x, n_gram){
  NGramTokenizer(x, Weka_control(min = n_gram, max = n_gram))
} 

get_corpus_freq <- function(my_corpus, n_gram = 1){
  ntoken <- function(x) ngram_Tokenizer(x, n_gram)
  dtm_matrix <- as.matrix(
    DocumentTermMatrix(my_corpus, control = list(tokenize = ntoken)))
  frequency <- colSums(dtm_matrix)
  frequency <- sort(frequency, decreasing = TRUE)
  return(frequency)
}

get_corpus_words <- function(my_corpus, n_gram = 1){
  return(names(get_corpus_freq(my_corpus, n_gram)))
}
```

#### 6.1 Display wordclouds

```{r}
display_wordcloud <-
  function(my_corpus,
           num = 100,
           n_gram = 1,
           palette = brewer.pal(8, "Dark2")) {
    frequency <- head(get_corpus_freq(my_corpus, n_gram), num)
    words <- names(frequency)
    my_title <- meta(my_corpus, "id")
    wordcloud(words, frequency, colors = palette)
    text(x = 0.5, y = 1, my_title)
  }

layout(matrix(c(1, 1, 2, 3), 2, 2, byrow = TRUE))
display_wordcloud(crps[1])  # blogs
display_wordcloud(crps[2])  # news
display_wordcloud(crps[3])  # twitter
```


#### 6.2 Plot frecuencies by word

```{r}
freq_df <- data.frame(head(get_corpus_freq(crps[1]), 15),
                      head(get_corpus_freq(crps[2]), 15),
                      head(get_corpus_freq(crps[3]), 15))
names(freq_df) <- list.files(sample_dir(LANGUAGE), pattern = "*.txt")
freq_df$names <- rownames(freq_df)
freq_df <- melt(freq_df, id.vars = 4)

ggplot(freq_df, aes(reorder(names, value), value)) +   
  geom_bar(aes(fill = variable), position = "dodge", stat = "identity") +
  coord_flip() +
  xlab("") +
  ylab("")
```

 
# 7. FULL SAMPLED CORPUS n-grams FRECUENCIES 


```{r}
head(get_corpus_freq(crps, n_gram = 1), 10)
head(get_corpus_freq(crps, n_gram = 2), 10)
head(get_corpus_freq(crps, n_gram = 3), 10)
head(get_corpus_freq(crps, n_gram = 4), 10)

ggplot_ngram <- function(my_corpus, n = 1, num = 30){
  df <- data.frame(head(get_corpus_freq(my_corpus, n_gram = n), num))
  names(df) <- "value"
  df$names <- rownames(df)
  my_color <- brewer.pal(8, "Dark2")[n]
  ggplot(df, aes(reorder(names, value), value)) +   
    geom_bar(stat = "identity", fill = my_color) +
    coord_flip() +
    xlab("") +
    ylab("") +
    theme_minimal()
}


```

```{r}
ggplot_ngram(crps, n = 1)
ggplot_ngram(crps, n = 2)
ggplot_ngram(crps, n = 3)
```



## Links:

